---
title: "Project proposal"
author: "CPL"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')
```

*For instructions on what each section should include, please see the [project page](https://idsed.digital/assessments/project/#proposal) on the course website. Remove this text when completing your proposal*.

## 1. Introduction

We got this dataset from Tidytuesday, and it was collected by JThomasmock from www.sports-reference.com. 

Each case is the athletes, and the variables are id, name, sex, age, height, weight, team, noc, games, year.

General research question:

1. Are younger athletes more likely to have a better performance than older athletes? 
 
2.Does the athletes from the host country tends to behave better by gaining more medals?

## 2. Data

```{r}
glimpse(olympics)
```

## 3. Data analysis plan

Are younger athletes more likely to have a better performance than older athletes? 
```{r}
population <- olympics %>%
  count(age)

medal_by_age <- olympics %>%
  na.omit(medal) %>%
  group_by(age) %>%
  mutate(n_medal = medal %in% c("Bronze", "Silver", "Gold")) %>%
  count(n_medal) %>%
  select(age, n) 

medal_by_age_2 <- medal_by_age %>%
  left_join(population, by = "age")

medal_by_age_2 %>%
  mutate(proportion = n.x/n.y) %>%
  ggplot(aes(x = age, y = proportion)) +
  geom_line()

```

Narrative: We think that the answer for the question is yes. The peak is at about 27~28 years old. The likelihood is gradually increasing from young age, have a peak at 27~27, and decreasing over age. But, the interesting thing is that there are 2 bumps at about 42 years old and 53 years. We want to find out why if we can.  We suppose there is the peak at 27~28 years old even though early twenties have better physical performance because early twenties have less experience for the Olympics, which is a big and making nervous competition, or have spent their time for the sports less than the older twenties. Anyway, it is generally true that the younger athletes are more likely to have a better performance than older athletes.



Which team tends to perform better in Winter Olympics than Summer Olympics? 
```{r}
total_medal_by_team <- olympics %>%
  group_by(team) %>%
  select(team, season, medal) %>%
  na.omit(medal) %>%
  mutate(n_medal = medal %in% c("Bronze", "Silver", "Gold")) %>%
  count(n_medal) %>%
  select(team, n)

winter_medal_by_team <- olympics %>%
  group_by(team) %>%
  select(team, season, medal) %>%
  filter(season == "Winter") %>%
  na.omit(medal) %>%
  mutate(n_medal = medal %in% c("Bronze", "Silver", "Gold")) %>%
  count(n_medal) %>%
  select(team, n)
  
winter_medal_by_team %>%
  left_join(total_medal_by_team, by = "team") %>%
  mutate(proportion = n.x/n.y) %>%
  arrange(desc(proportion)) %>%
  filter(n.y >=100) %>%
  rename(winter_medal = n.x,
         all_medal = n.y) %>%
  mutate(team = case_when(
    team == "United States-1" ~ "United States",
    TRUE ~ team
  ))


```

Narrative: We think it would be more clear if we demonstrate the answer as a table, so we did. We calculated the proportion by dividing the number of medal each team got on winter by all the medal they got from both Winter Olympics and Summer Olympics. The first 5 countries on this table are mostly the countries who have cold weather and mountains, so we think it is reasonable for them to be behaving better on winter. As we can see from the table, Untied States tends to perform better on Winter Olympics because they have got 63 medals on winter out of 101. The proportion is 68.3% high. 


Specific questions: 
1. The number of athletes per age and see what kind of sports that younger athletes are good at using ggplot.
```{r}
olympics %>%
  group_by(age) %>%
  mutate(n_medal = medal %in% c("Bronze", "Silver", "Gold")) %>%
  count(n_medal) %>%
  select(age, n)
```

```{r}
olympics %>%
  group_by(sport) %>%
  na.omit(medal) %>%
  mutate(n_medal = medal %in% c("Bronze", "Silver", "Gold")) %>%
  filter(sport %in% c("Athletics","Gymnastics","Swimming","Shooting",		"Cycling", "Fencing", "Rowing",	"Cross Country Skiing",	"Alpine Skiing", "Wrestling", "Football", "Sailing")) %>%
  ggplot(aes(x = sport, y = age, colour = age)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 30))

olympics %>%
  group_by(sport) %>%
  na.omit(medal) %>%
  mutate(n_medal = medal %in% c("Bronze", "Silver", "Gold")) %>%
  ggplot(aes(y = sport, x = age, colour = sport)) +
  geom_boxplot() +
  theme(axis.text.y = element_text(size = 6), legend.position = "none")
  
olympics %>%
  count(sport, sort = TRUE)

```

2. Which sports cause the two spikes around the age 46 and 52?

These are the enlargement of the previous graphs from the age 44 to 55.
```{r}
population <- olympics %>%
  count(age)

medal_by_age <- olympics %>%
  na.omit(medal) %>%
  group_by(age) %>%
  mutate(n_medal = medal %in% c("Bronze", "Silver", "Gold")) %>%
  count(n_medal) %>%
  select(age, n) 

medal_by_age_2 <- medal_by_age %>%
  left_join(population, by = "age")

medal_by_age_2 %>%
  mutate(proportion = n.x/n.y) %>%
  ggplot(aes(x = age, y = proportion)) +
  geom_line() +
  xlim(44,55)
```

```{r}
olympics %>%
  group_by(age) %>%
  mutate(n_medal = medal %in% c("Bronze", "Silver", "Gold")) %>%
  count(age) %>%
  select(age, n) %>%
  ggplot(aes(x = age, y = n)) +
  geom_line() +
  xlim(44,55) +
  ylim(0,1000)
```



This is the number of medals the athletes (from age 44 to 48) got by sport.
```{r}
olympics %>%
  filter(age > 44, age < 48 ) %>%
  na.omit(medal) %>%
  count(sport, sort = TRUE)
```
This is the number of medals the athletes (from age 50 to 54) got by sport.
```{r}
olympics %>%
  filter(age > 50, age < 54 ) %>%
  na.omit(medal) %>%
  count(sport, sort = TRUE)
```



```{r}
olympics %>%
  filter(sport %in% c("Equestrianism", "Sailing", "Shooting")) %>%
  na.omit(medal) %>%
  group_by(age) %>%
  mutate(n_medal = medal %in% c("Bronze", "Silver", "Gold")) %>%
  ggplot(aes(x = age, y = sport, colour = age)) +
  geom_boxplot()
```

According to the previous two tibbles, the top three number of medals are shooting, sailing and equestrianism. The box plot shows the range and the density for each three of them. We can see that for equestrianism, the median and the range of age are significantly higher than the other two. 
(in the presentation, maybe we can use the boxplot for the overall sports again here to make conclusions about the spikes)
Also from the box plot in the previous question, equestrianism has the highest median among all the sports. Although shooting and sailing do not have such high median, the interquartile range for these two sports are outstandingly higher than the others.

3. Make a model related to the height and the probability of wining medals.
```{r}
#load the modelling pacakage
library(tidymodels)
```
```{r}
#model_olympics <- olympics %>%
#  group_by(medal) %>%
#  na.omit(medal) %>%
#  count(medal) %>%
#  mutate(proportion_medal = (10148+10167+9866)/271116)
#model_olympics


#linear_reg()%>%
 # set_engine("lm")%>%
  #fit(height ~ proportion_medal, data = model_olympics)

```

```{r}

set.seed(1369)
# Split into 80% and 20% 
olympics_split <- initial_split(olympics, prop = 0.80)
train_data <- training(olympics_split)
test_data <- testing(olympics_split) 

```



