---
title: "Project proposal"
author: "CPL"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')
```

*For instructions on what each section should include, please see the [project page](https://idsed.digital/assessments/project/#proposal) on the course website. Remove this text when completing your proposal*.

## 1. Introduction

We got this dataset from Tidytuesday, and it was collected by JThomasmock from www.sports-reference.com. 

Each case is the athletes, and the variables are id, name, sex, age, height, weight, team, noc, games, year.

General research question:
1. What is the correlation between the height of the athletes and the medals, if there is any?   

2. Are younger athletes more likely to have a better performance than older athletes? 

3. Which team tends to perform better in Winter Olympics than Summer Olympics? 

4. Is the percentage of females attending the Olympics increasing over years? 

5.Does the athletes from the host country tends to behave better by gaining more medals?

## 2. Data

There are 16,598 observations and 11 variables. 
```{r}
glimpse(olympics)
```

## 3. Data analysis plan



